---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "out/README_"
)
```

<!-- This document generates README.md upon Knitting in RStudio -->

# funDecomp 

Code for the paper 
*Unemployment Volatility: When Workers Pay Costs upon Accepting Jobs*.
The paper is included in the top of the directory in 
`unemployment_volatility_when_workers_pay_costs.pdf`.

Here's the 100-word abstract:

> Hiring workers is costly.
> Firms' costs
> reduce resources that can go to recruitment
> and
> amplify how unemployment responds to changes in productivity.
> Workers also incur up-front costs. 
> Examples include moving expenses and regulatory fees.
> Workers' costs
> lessen unemployment volatility and 
> leave resources available for recruitment unchanged. 
> Their influence is bounded by properties of a matching function.
> Using adjusted data on job finding,
> I estimate a bound that ascribes limited influence.
> The results demonstrate that workers' costs 
> affect outcomes
> (firms threaten workers with paying the fixed costs again if negotiations fail),
> but their influence on volatility is less than firms' costs.  

<!-- > When a firm hires a worker, -->
<!-- >   adding the new hire to payroll is costly. -->
<!-- >   These costs -->
<!-- >   reduce the amount of resources that can go to recruiting workers -->
<!-- >   and affects -->
<!-- >   amplify how unemployment responds to changes in productivity. -->
<!-- >   Workers also incur up-front costs upon accepting jobs. -->
<!-- >   Examples include moving expenses and regulatory fees. -->
<!-- >   I establish that workers' costs -->
<!-- >   lessen the response of unemployment to productivity changes and -->
<!-- >   do not subtract from resources available for recruitment. -->
<!-- >   The influence of workers' costs is bounded by properties of a matching function, -->
<!-- >   which describes how -->
<!-- >   job openings and unemployment produce hires. -->
<!-- >   Using data on job finding that are adjusted -->
<!-- >   for workers' transitions between employment and unemployment and  -->
<!-- >   for how the Job Openings and Labor Turnover Survey records hires, -->
<!-- >   I estimate a bound that ascribes limited influence to workers' costs.  -->
<!-- >   The results demonstrate that costs paid by workers upon accepting jobs -->
<!-- >   affect outcomes in the labor market -->
<!-- >   (firms threaten workers with paying the up-front costs again if wage -->
<!-- >   negotiations fail), -->
<!-- >   but their influence on volatility is less important than firms' costs. -->

## Accessing adjusted data on job-finding and job-separation rates ##

Here is code that 
collects relevant data and 
plots the adjusted and unadjusted rates of job finding.

Here's code that 
```{r data, message=FALSE, warning=FALSE}
library(tidyverse) # for useful R functionality
library(here)      # for accessing the data

dat <- read_csv(here("out", "dat_02-adjust-transition-rates.csv"))

dat_useful <- dat |> 
  select(date, ends_with("_crude"), find, sep)

# Save the useful data
fout <- here("out", "dat_funDecomp.csv")
write_csv(dat_useful, file = fout)

head(dat_useful)
```

The paper contains a detailed account of the variables.
Briefly, the variables are:

 + `date` : monthly date
 + `find_crude` : number of hires divided by the unemployment level
 + `find` : rate of job finding adjusted to account for workers' transitions and hires recording all hires, even hires fired before the end of the month
 + `sep_crude` : unadjusted separation rate
 + `sep` : adjusted separation rate

Here is a figure that uses the data:

```{r plot-find, fig.alt = "Plot of adjusted and unadjusted rates of job finding.", message=FALSE, warning=FALSE}
datfig <- read_csv(here("out", "dat_funDecomp.csv")) 

# Put the data into long form for plotting
datfig <- datfig |> 
  select(date, starts_with("find")) |> 
  pivot_longer(starts_with("find")) 
  
csub_blue <- rgb(0, 26, 112, maxColorValue = 255)

ggplot(data = datfig) +
  geom_hline(yintercept = 1.0, color = "maroon", linetype = "dotted") +
  geom_line(mapping = aes(x = date, y = value, color = name, linetype = name), size = 0.9) +
  scale_color_manual(values = c("find_crude" = "black", "find" = csub_blue)) +
  scale_linetype_manual(values = c("find" = "solid", "find_crude" = "dashed")) +  
  labs(x = "", y = "", title = "Monthly probability of finding a job", 
       color = "Series", linetype = "Series") +
  theme_light()  
```

Code that produces better-looking figures is contained in `cde/`.

## Overview of Code ##

Code that replicates that project is contained in `cde/`.
The scripts are prefixed with the order in which they need to be run.

  * `01-get-FRED-data.R`: Retrieves data from [FRED](https://fred.stlouisfed.org/).
  * `02-adjust-transition-rates.R`: Uses the FRED data to adjust rates of 
  job finding and job separation to account for 
      * workers' transitions between employment and unemployment and
      * how the Job Openings and Labor Turnover Survey (JOLTS) records hires---all
	  hires for a month are recorded, 
	  even those who separate before the month ends.
  * `03-est-matching.R`: Estimates the nonlinear mathching technology,
	$M \left( u, v \right) = \mu \frac{uv}{\left( u^{\gamma} + v^{\gamma} \right)^{1/\gamma}}$.   Various plots are produced showing how 
  the Beveridge curve shifted since JOLTS data on vacancies 
  became available and the fit produced by the estimated model.
  * `04-plot-elasticity.R`: Plots the elasticity of matching
  with respect to unemployment, $\eta_{M,u}$, and
  the upper bound for $\Upsilon$,
  $\max \left( \frac{1}{\eta_{M,u}}, \frac{1}{1-\eta_{M,u}} \right)$.
  The factor $\Upsilon$ is one of the factors 
  in [Ljungqvist and Sargent's (2017)](https://www.aeaweb.org/articles?id=10.1257/aer.20150233) fundamental decomposition of the 
  elasticity of market tightness with respect to productivity, $\eta_{\theta,y}$.
  The code also generates the tables in the text.
  * `05-plot-replacement-rates.R`: Plots two measures of how unemployment-insurance benefits replace claimants' labor income.

## Note about Viewing a Function's Documentation

After running an script,
it is possible to view a function's documentation using `?my_function`.
For example,
`docstring::docstring(compute_c)` returns the documentation of the function `compute_c`.

This feature is accomplished with R's docstring package.

## Citation

Ryan, Rich. (2024) Unemployment volatility: When Workers Pay Costs upon Accepting Jobs. _International Journal of Economic Theory_, 20, 303--333. [https://doi.org/10.1111/ijet.12405](https://doi.org/10.1111/ijet.12405)

## License

 <p xmlns:cc="http://creativecommons.org/ns#" xmlns:dct="http://purl.org/dc/terms/"><a property="dct:title" rel="cc:attributionURL" href="https://github.com/richryan/funDecomp">Replication Material for "Unemployment Volatility: When Workers Pay Costs upon Accepting Jobs"</a> by <a rel="cc:attributionURL dct:creator" property="cc:attributionName" href="https://richryan.github.io/">Rich Ryan</a> is licensed under <a href="https://creativecommons.org/licenses/by-sa/4.0/?ref=chooser-v1" target="_blank" rel="license noopener noreferrer" style="display:inline-block;">CC BY-SA 4.0<img style="height:22px!important;margin-left:3px;vertical-align:text-bottom;" src="https://mirrors.creativecommons.org/presskit/icons/cc.svg?ref=chooser-v1" alt=""><img style="height:22px!important;margin-left:3px;vertical-align:text-bottom;" src="https://mirrors.creativecommons.org/presskit/icons/by.svg?ref=chooser-v1" alt=""><img style="height:22px!important;margin-left:3px;vertical-align:text-bottom;" src="https://mirrors.creativecommons.org/presskit/icons/sa.svg?ref=chooser-v1" alt=""></a></p> 